dplyr In a Nutsehll
===

```{r setup, include=FALSE, echo=FALSE}
opts_chunk$set(comment=NA)
```

# 8 dplyr Functions to Rule the World

### Speedy Table

`tbl_df`


### The 5 Guys + 1

1. `filter`
2. `select`
3. `mutate`
4. `group_by`
5. `summarise`
6. `arrange`

### Chaining

`%.%`

# Relating the Functions

### Speedy Table 

`tbl_df` works like similar to `data.table` 

### Relating the 5 Guys + 1 to base R

List of dplyr functions and the base functions they're related to:

Base Function    | dplyr Function(s) | Special Powers
-----------------|-------------------|-----------------------------
subset           |  filter & select  | filter rows & select columns
transform        |  mutate           | operate with columns not yet created
split            |  group_by         | splits without cutting
lapply + do.call |  summarise        | apply and join in a single bound
order + with     |  arrange          | "I only have to specify dataframe once?"

### Chaining

`%*%` Do you know ggplot2's `+`?  Same idea.  

*Basically previous input in chain supplied as argument 1 to function on right side.*

# Demos

```{r}
library(dplyr)

mtcars2 <- tbl_df(mtcars)

mtcars2 %.%
    group_by(cyl) %.%
    summarise(md=mean(disp), mh=mean(hp), mdh=mean(disp + hp))
mtcars2 %.%
    group_by(cyl, gear) %.%
    summarise(md=mean(disp), mh=mean(hp), mdh=mean(disp + hp)) %.%
    arrange(-cyl, -gear)
## Use `%.%` with base functions too!!!
mtcars2 %.%
    group_by(cyl, gear) %.%
    summarise(md=mean(disp), mh=mean(hp), mdh=mean(disp + hp)) %.%
    arrange(-cyl, -gear) %.%
	head()
```



